---
title: "Machine Learning Project; B. Jalali; April 2015"
output: html_document
---

# 1. Executive Summary
We use random forest model to predict the manner of 20 observations on doing 
weight lifting exercise. In this experiment six participants were asked to perform 
Dumbbell Biceps Curl exercise in five different fashions (Classes A, B, C, D and E),
of which only class A is the correct way. 

Cross validation on 25% of the training data reveals 99% out of sample accuracy. 
OOB estimate of error rate is about 0.6%. Further, using **importance(model)** we see that **roll_belt** and **yaw_belt** are among the most important predictors.

We successfully predict all 20 test observations correctly!

# 2. Reading and Preprocessing the data 
```{r,echo=TRUE, , message=FALSE}
library(caret); library(kernlab); library(randomForest)
```

## Read the training csv file 
```{r,echo=TRUE, , message=FALSE}
data_training <- read.csv("pml-training.csv", na.strings= c("NA",""," "))
```

## Cleaning the data by removing columns with NAs
```{r,echo=TRUE, , message=FALSE}
data_training_NA <- apply(data_training, 2, function(x) {sum(is.na(x))})
data_training_clean <- data_training[, which(data_training_NA == 0)]

# Final preprocessing by removing not useful 1-7th columns
data_training_clean <- data_training_clean[8:length(data_training_clean)]
```

# 3. Split the cleaned training data into two sub groups: training and testing for cross validation
```{r,echo=TRUE, , message=FALSE}
# set a seed number
set.seed(2015)

inTrain <- createDataPartition(y = data_training_clean$classe, p = 0.75, list = FALSE)
training <- data_training_clean[inTrain, ]
testing_cv <- data_training_clean[-inTrain, ]
```

# 4a. Fit a model to predict the classe using all the rest of the variables as predictors
```{r,echo=TRUE, , message=FALSE}
model <- randomForest(classe ~ ., data = training, ntree=101)
model
```

# 4b. Cross validating the model using the remaining 25% of the training data
```{r,echo=TRUE, , message=FALSE}
predict_cv <- predict(model, testing_cv)
confusionMatrix(testing_cv$classe, predict_cv)
```

We obtained 0.99 accuracy in our modeling of sub-training data and using the 
sub-training test group. 

# 5. Final prediction on the pml-test testing data after the same preprocessing steps
```{r,echo=TRUE, , message=FALSE}
data_test <- read.csv("pml-testing.csv", na.strings= c("NA",""," "))
data_test_NA <- apply(data_test, 2, function(x) {sum(is.na(x))})
data_test_clean <- data_test[, which(data_test_NA == 0)]
data_test_clean <- data_test_clean[8:length(data_test_clean)]

# Predict the classes of the test data 
predict_test <- predict(model, data_test_clean)
```

# 6. Writing predicted answers in files
```{r,echo=TRUE, , message=FALSE}
answers <- as.character(predict(model, data_test_clean))
pml_write_files = function(x){
        n = length(x)
        for(i in 1:n){
                filename = paste0("problem_id_",i,".txt")
                write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
        }
}
pml_write_files(answers)
```
